An Evolutionary Algorithm for the Multiobjective Risk-Equity Constrained Routing Problem

### 2011-11-28 DAW_hazmatHypE/trunk/paper/hazmatRouting_EvolOptim.tex
decided on final instances

As a test case for all the experiments, we chose to use a real-world case study based on the work of \citet{bianco09}. The network used here consists of 311 nodes and 441 arcs and represents the main transport roads in Italy's Lazio region which, amongst other cities, includes the capital Rome. We distinguish between instances with a different number of commodities (between 2 and 4) and with different numbers of trucks, see Table~\ref{tab:instances}. While the costs $c_{ij}$ are proportional to the real distance between the two locations $i$ and $j$, the risks $r_{ij}^c$ for each arc $(i,j) \in A$ and commodity $c \in C$ have been computed as a product of the arc's length, the actual population density in the area of the arc and a constant which depends on the commodity, for details see \citep{bianco09} and Table~\ref{tab:instances}. Solely one instance is created differently: to show the impact of correlation between the costs and the risks, we created the instance lazio$_{\text ncorr}$\_4\_1 with the same graph and risks as for the instance lazio\_4\_1 but with randomly permutated costs where each occuring cost value has been divided by a random value between 0 and 5000.

\begin{table}
\caption{\label{tab:instances} Parameters of the instances used in the experimental analyses.}
\begin{tabular}{lllll}
	\toprule
	name & \begin{minipage}{1cm}\#com\-mod\-i\-ties\end{minipage} & \#trucks & cost range & risk range \\ \midrule
	lazio\_2\_1 & 2 & \small 20+19         & \multirow{2}{*}{\small [0.25, 41]} & \multirow{2}{*}{\begin{minipage}[t]{3.7cm} \small [762, 274628]\\ x [610.18, 219702.6]\end{minipage}}\\
	\cmidrule{1 - 3}
	lazio\_2\_2 & 2 & \small 23+29         & & \\
	\midrule
	lazio\_3 & 3 & \small 25+16+28      & \small [0.25, 41] & \begin{minipage}[t]{3.7cm} \small [762, 274628]\\ x [610.18, 219702.6]\\ x [1120.352, 439405.2]\end{minipage}\\
	\midrule
	lazio\_4 & 4 & \small 21+3+30+5     & \begin{minipage}[t]{2cm} \small [0.25, 41]\\ \end{minipage} & \multirow{2}{*}{\begin{minipage}[t]{3.7cm} \small [762, 274628]\\ x [610.18, 219702.6]\\ x [1120.352, 439405.2]\\ x [762, 274628]\end{minipage}}\\
	\cmidrule{1 - 4}
	uncorrelated\_4 & 4 & \small 21+3+30+5     & \begin{minipage}[t]{2cm} \small [0.0109239, 3.93701]  \end{minipage} & \\
	\bottomrule
\end{tabular}
\end{table}

@@

\subsection{Experimental Settings}
Except otherwise stated, we used the following parameter setting. We perform 10 independent runs of HypE for 2000 generations with a population size of 100. The reference point for the hypervolume indicator was chosen as in Table~\ref{tab:referencepoints} according to the initial objective function values, such that all initial solutions have a positive contribution to the hypervolume indicator\footnote{Note that the reason for using the same values in all objectives is the used PISA implementation of HypE.}. In all cases, the hypervolume computation is exact.

\begin{table}
\centering
\caption{\label{tab:referencepoints} Used reference points within the environmental selection of HypE}
\begin{tabular}{lr}
	\toprule
	instance & reference point\\ \midrule
	\# commodities: 2 & $(10^8, 10^8, 10^8)$\\
	\# commodities: 3 & $(10^8, 10^8, 10^8)$\\
	\# commodities: 4 & $(10^9, 10^9, 10^9)$\\
	\bottomrule
\end{tabular}
\end{table}

\subsection{Differences Between Instances}

@@

\subsection{The Feasibility of the EMO Approach}
\begin{itemize}
	\item of course no guarantee that the solutions are Pareto-optimal
	\item however, in reasonable time (e.g. ??hrs for 10 runs with 1000 generations each) results with several non-dominated solutions the decision maker can interpret and with which she can get an idea about the trade-offs between the objectives
	\item if DM wants good results fast, cost-efficient initialization is favourable, otherwise, random initialization gives better spread of the solutions
	\item if needed, approach can be combined easily with preference articulation techniques such as the weighted hypervolume indicator which allows to steer the search towards user-defined regions of the objective space (show 1 plot with W-HypE).
	\item approach is very general and can easily deal with slightly modified problems: e.g. different truck loads (only the objective function code needs to be changed)
	\item it needs to be mentioned that for the tested instances with 311 nodes and 441 arcs, the variation operator of the evolutionary algorithm was able to construct the shortest paths from source to destination in some of the runs for some of the truck paths within the first 100,000 function evaluations. This corresponds to performing random walks of lengths around 20 nodes.
\end{itemize}

>>>>

As a test case for all the experiments, we chose to use a real-world case study based on the work of \citet{bianco09}. The network used here consists of 311 nodes and 441 arcs and represents the main transport roads in Italy's Lazio region which, amongst other cities, includes the capital Rome. We distinguish between instances with a different number of commodities (between 2 and 4) and with different numbers of trucks, see Table~\ref{tab:instances}. While the costs $c_{ij}$ are proportional to the real distance between the two locations $i$ and $j$, the risks $r_{ij}^c$ for each arc $(i,j) \in A$ and commodity $c \in C$ have been computed as a product of the arc's length, the actual population density in the area of the arc and a constant which depends on the commodity, for details see \citep{bianco09} and Table~\ref{tab:instances}. Solely one instance is created differently: to show the impact of correlation between the costs and the risks, we created the instance uncorrelated\_4 with the same graph and risks as for the instance lazio\_4\_1 but with randomly permutated costs where each occuring cost value has been divided by a random value between 0 and 5000.

\begin{table}
\caption{\label{tab:instances} Parameters of the instances used in the experimental analyses with $|C|$ being the number of different commodities.}\vspace{1em}
\begin{tabular}{lllll}
	\toprule
	name & $|C|$ & \#trucks & cost range & risk range \\ \midrule
	lazio\_2 & 2 & \small 25+16         & \small [0.25, 41] & \begin{minipage}[c]{3.8cm} \small [762, 274628]\\ $\times$ [610.18, 219702.6]\end{minipage}\\
	\midrule
	lazio\_3 & 3 & \small 25+16+28      & \small [0.25, 41] & \begin{minipage}[c]{3.8cm} \small [762, 274628]\\ $\times$ [610.18, 219702.6]\\ $\times$ [1120.352, 439405.2]\end{minipage}\\
	\midrule
	lazio\_4\_1 & 4 & \small 21+3+30+5     &  & \multirow{4}{*}{\begin{minipage}[c]{3.8cm} \hspace{1pt}\\ \\ \small [762, 274628]\\ $\times$ [610.18, 219702.6]\\ $\times$ [1120.352, 439405.2]\\ $\times$ [762, 274628]\end{minipage}}\\
	\cmidrule{1 - 3}
	lazio\_4\_2 & 4 & \small 21+3+30+6     & \small [0.25, 41] & \\
	\cmidrule{1 - 3}
	lazio\_4\_3 & 4 & \small 7+1+10+2     &  & \\
	\cmidrule{1 - 4}
	uncorrelated\_4 & 4 & \small 21+3+30+5     & \begin{minipage}[t]{2cm} \small [0.0109239, 3.93701]  \end{minipage} & \\
	\bottomrule
\end{tabular}
\end{table}

@@

\subsection{Experimental Settings}
Except otherwise stated, we used the following parameter setting. We perform 10 independent runs of HypE for 2000 generations with a population size of 50, i.e., each algorithm run has a budget of 100,000 function evaluations. The reference point for the hypervolume indicator was chosen as $(10^8, 10^8, 10^8)$ for the instances with 2 and 3 commodities and as $(10^9, 10^9, 10^9)$ for all instances with 4 commodities. The reference point has been chosen according to the objective function values of the random initial populations, such that all initial solutions have a positive contribution to the hypervolume indicator. Note that the reason for using the same values in all objectives is the used PISA implementation of HypE. In all cases, the hypervolume computation is exact.

\begin{table}
\centering
\caption{\label{tab:referencepoints} Used reference points within the environmental selection of HypE}
\vspace{1em}
\begin{tabular}{lr}
	\toprule
	instance & reference point\\ \midrule
	\# commodities: 2 & $(10^8, 10^8, 10^8)$\\
	\# commodities: 3 & $(10^8, 10^8, 10^8)$\\
	\# commodities: 4 & $(10^9, 10^9, 10^9)$\\
	\bottomrule
\end{tabular}
\end{table}

\subsection{Differences Between Instances}
The basis for our experiments are three instances of the lazio network with 2 (lazio\_2), 3 (lazio\_3), and 4 commodities (lazio\_4\_1) which differ also differ in the number of trucks for the commodities. Whereas the number of trucks for lazio\_3 and lazio\_4\_1 have been chosen randomly, lazio\_2 has the same number of trucks than the lazio\_3 instance for the first two commodities.

\TODO{show results and talk about them}

@@

\subsection{The Feasibility of the EMO Approach}
Due to their heuristic nature, evolutionary multiobjective optimization algorithms like the one presented in this study cannot guarantee that the solutions found are actually Pareto-optimal. However, as we have seen above, the proposed algorithm is able to produce a set of non-dominated solutions in reasonable time which are close to Pareto-optimal solutions found by exact algorithms. Such sets of non-dominated solutions can be a starting point for a decision maker to investigate the trade-offs between the objectives and to decide in which objective space region interesting Pareto-optimal solutions are which can be explored further with more involved algorithms---be it exact algorithms or specialized evolutionary algorithms such as the mentioned W-HypE variant. In the above example instances, it turned out that total risk and cost are indeed more or less fully correlated in the original problem formulation or only loosely (with a larger trade-off between cost and total risk) when the individual arcs' risks and costs are.

Regarding the two initialization schemes, it turned out that in combination with the (local and simple) variation operator, the cost-efficient initialization does not allow the algorithm to find additional solutions besides the initial one and that, in particular if the cost and total risk objectives are correlated, the random initialization is beneficial. Using other initializations as well as additional variation operators from the literature is easily possible with the proposed algorithm but their investigation is left for future research\footnote{For example, the

On the other hand, the proposed variation operator can be changed towards a larger neighborhood by allowing two or more nodes to be added or deleted from each truck path at a time---resulting in more exploration than exploitation when compared to the current variation operator. Also well-designed crossover operators are expected to be beneficial to the problem when, for example, good sub paths can be exchanged between solution paris.
}.

\begin{itemize}

	\item approach is very general and can easily deal with slightly modified problems: e.g. different truck loads (only the objective function code needs to be changed)
	\item it needs to be mentioned that for the tested instances with 311 nodes and 441 arcs, the variation operator of the evolutionary algorithm was able to construct the shortest paths from source to destination in some of the runs for some of the truck paths within the first 100,000 function evaluations. This corresponds to performing random walks of lengths around 20 nodes.
\end{itemize}
